---
title: "Tidy Tuesday 1"
author: "Kyle Bosworth"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../outputs/")
knitr::opts_chunk$set(dev = "png", 
                      fig.path = "../outputs/",
                      fig.width = 7,
                      fig.height = 5)

```

### Load Libraries
```{r}
library(here)
library(tidyverse)
library(gganimate)
library(gifski)
library(transformr)
```

### Read in data: Monster Movies
```{r}
monster_movie_genres <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-10-29/monster_movie_genres.csv')
monster_movies <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-10-29/monster_movies.csv')

glimpse(monster_movies)
```



### Goal/skill that I have not done yet: Adding animation to my plots

Which year had the highest rated horror movie?
Data clean up
- Use subset movie, year, average rating
- top movie of each year then arrange in terms of top 10 highest rated moview overall
-Letʻs try to display these movies to pop up in ascending years (past to present) based on their rating

#### Data cleanup: filter and remove NAs
```{r}
Horror_movies <- monster_movies %>%
  filter(str_detect(genres, "Horror")) %>% #detecting string patter in this case (genres, Horror)
  select(original_title, year, average_rating) %>% #from the horror genres we only want the original title, year, and average rating
  filter(!is.na(average_rating) & !is.na(year)) # we want to get rid of all NAs in the year and average rating columns

head(Horror_movies) #chk
```

#### Data cleanup: top 10 arrange 1-10
```{r}
top10_by_year <- Horror_movies %>% 
  group_by(year) %>% #top rated movie for each year
  slice_max(average_rating, n=1) %>%
  ungroup()

top10 <- top10_by_year %>% #top rated movie overall
  slice_max(average_rating, n=10) %>%
  arrange(desc(average_rating))
```

#### Make a plot
```{r}
ggplot(top10,
       aes(x = reorder(original_title, average_rating), #reorder(titles by rating)
           y = average_rating,
           fill = as.factor(year))) + #use as.factor for discrete colors
  geom_col() +
  geom_text(aes(label = paste(original_title, " (", year, ")", sep = "")), 
            hjust = 1, nudge_y = -0.1, color = "white") +
  coord_flip() +
  scale_fill_viridis_d() +
  labs(title = "Top10 Highest Rated Horror Movies",
       x = NULL,
       y = "Average Rating",
       fill = "Year") +
  theme_minimal() +
   theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        axis.text.y = element_blank())
```
#### Addmy animation to the plot
```{r}
animated_plot <- ggplot(top10, aes(x = reorder(original_title, average_rating), y = average_rating, fill = as.factor(year))) +
  geom_col() +
  geom_text(aes(label = paste(original_title, " (", year, ")", sep = "")), #Use paste() to combine two variables into one lable, formatted: ʻmovie title(year)ʻ
            hjust = 1, 
            nudge_y = -0.1, 
            color = "white") + 
  coord_flip() +
  scale_fill_viridis_d() +
  labs(title = "Top 10 Highest-Rated Horror Movies",
       subtitle = "Year: {frame_along}",  #add dynamic subtitle ov year progression
       x = NULL,
       y = "Average Rating",
       fill = "Year") +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        axis.text.y = element_blank()) +
  transition_reveal(year)  #animates the plot by fading-in scale bars by year

```

#### Display GIF and save

```{r create-and-display-gif, echo=FALSE, fig.cap="Top 10 Highest-Rated Horror Movies"}
anim_save("./outputs/top_10_horror_movies.gif", animated_plot)

gif_path <- here::here("Tidy Tuesdays", "TT1", "outputs", "top_10_horror_movies.gif")
anim_save(gif_path, animated_plot)
knitr::include_graphics(gif_path)
```