---
title: "week6HW"
author: "Kyle Bosworth"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    css: snazzy.css

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 10, 
                      fig.height = 6, 
                      out.width = "100%", 
                      fig.retina = 3,
                      fig.align = "center")
```

### Load libraries
```{r}
library(tidyverse)
library(here)
library(ggplot2)
library(kableExtra)
library(prettydoc)
```

### Load and clean up data
```{r}
maunaluaSGD<-read.csv(here("week6","data","chemicaldata_maunalua.csv"))
#Define the zones in order
zone_order<- c("Near Spring","Channel","Diffuse","Transition","Ambient","Offshore")
#create as factor
maunaluaSGD <- maunaluaSGD %>%
  drop_na(percent_sgd, Salinity, Zone, Season) %>%
  mutate(Zone = factor(Zone, levels = zone_order))
```

### Inspect the data
```{r}
glimpse(maunaluaSGD)
```

### _What is the relationship between percent Submarine Groud Water Discharge (SGD) and salinity across different coastal zones, and how does this relationship vary by season?_

### Figure 1. Relationship between Percent SGD and Salinity Across Coastal Zones in Fall and Spring
```{r}
ggplot(maunaluaSGD,
       aes(x = percent_sgd,
           y = Salinity,
           color = Zone,
           shape = Season)) + 
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, 
              aes(group = 1, 
                  color = "black", 
                  linetype = "dashed")) +
  scale_color_viridis_d() +
    facet_wrap(~ Season, scales = "free") +
    labs(title = "Percent SGD and Salinity",
         subtitle = "[Accross different zones and seasons in Maunalua Bay]",
         x = "Percent SGD",
         y = "Salinity (ppt)",
         color = "Zone",
         Shape = "Season") +
    theme_minimal() +
    theme(legend.position = "right",
          plot.title = element_text(hjust = 0.5, face = "bold"),
          plot.subtitle = element_text(hjust = 0.5, face = "italic"))
```

### Table 1. Mean Salinity and SGD (Fall)
```{r}
fall_table<-maunaluaSGD %>%
  filter(Season == "FALL") %>%
  group_by(Zone) %>% 
  summarise(
    mean_salinity = mean(Salinity),
    mean_percent_sgd = mean(percent_sgd)) %>%
  arrange(factor(Zone, levels = zone_order))

kable(fall_table, caption = "Table 1. Mean Salinity by Zone (Fall)") %>%
  kable_styling(bootstrap_options = c("striped",
                                      "hover",
                                      "condensed"), full_width = F)
```

### Table 2. Mean Salinity and SGD (SPRING)
```{r}
spring_table<-maunaluaSGD %>%
  filter(Season == "SPRING") %>%
  group_by(Zone) %>%
  summarise(
    mean_salinity = mean(Salinity),
    mean_percent_sgd = mean(percent_sgd)) %>%
  arrange(factor(Zone, levels = zone_order))

kable(fall_table, caption = "Table 2. Mean Salinity and SGD by Zone (SPRING)") %>%
  kable_styling(bootstrap_options = c("striped", 
                                      "hover", 
                                      "condensed"), full_width = F)
```


