---
title: "Week 13 Class notes for loop"
author: "Kyle Bosworth"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../outputs/")
knitr::opts_chunk$set(dev = "png", 
                      fig.path = "../outputs/",
                      fig.width = 7,
                      fig.height = 5)
```

### Load Libraries
```{r}
library(here)
library(tidyverse)
```


### For loops: 
## for loops are one way to run an iterative process (do something over and over again). However, with the tidyverse you can avoid needing loops for almost any basic coding needs. Where it does come in handy is in population modeling and Bayesian modeling, for example. Also, it is integral to programming and is happening "under the hood" whether you use them or not. 


##### or loops have two major parts: an indexing statement and a command (or set of commands) to repeat. The coding is in base R. 

####The command looks like this:for(index in sequence){command to repeat}

```{r}

print(paste("This year is",2000))

#put it in a for loop

years <- c(2015:2021)
for (i in years) { # set up the for loop where i is the index
  print(paste("The year is", i)) # loop over i
}

```
### What we just did printed something over and over, but it did not save it anywhere. Let's say you want to save a new vector with all the years. To do this we need to pre-allocate space and tell R where it is going to be saved. Create an empty dataframe called year_data with columns for year and year_name .

```{r}

#pre-allocate space for the for loop
# empty matrix that is as long as the years loop

year_data <- tibble(year = rep(NA, length(years)), # column name for year
                    year_name = rep(NA, length(years))) # column name for the year name
year_data

```

### Add the for loop

One line at a time. Let's first add in the column that is going to have all the names in it. Notice that I added an index i in the column name. I also am having the index go from 1:length(years), which is 1:7. I use length() because it allows me to change the number of years in the vector without having to change the for loop.
```{r}

for (i in 1:length(years)){ # set up the for loop where i is the index
  year_data$year_name[i]<-paste("The year is", years[i]) # loop over i
}
year_data


```


### Fill in you column (use for respirometry)

```{r}
for (i in 1:length(years)){ # set up the for loop where i is the index
  year_data$year_name[i]<-paste("The year is", years[i]) # loop over year name
  year_data$year[i]<-years[i] # loop over year
}
year_data

```

### testdata: conddata
```{r}
testdata<-read_csv(here("week13", "data", "cond_data", "011521_CT316_1pcal.csv"))

#we wnat to list a directory for our files

#### point to the location on the computer of the folder
CondPath<-here("week13", "data", "cond_data")

# list all the files in that path with a specific pattern
# In this case we are looking for everything that has a .csv in the filename
# you can use regex to be more specific if you are looking for certain patterns in filenames
files <- dir(path = CondPath,pattern = ".csv")
files


```

### BC we are doing a for loop, we need to pre-allocate space for the loop
Let's calculate the mean temperature and salinity from each file and save it, one file at a time
```{r}
# pre-allocate space
# make an empty dataframe that has one row for each file and 3 columns
 cond_data<-tibble(filename =  rep(NA, length(files)),  # column name for year
                   mean_temp = rep(NA, length(files)), # column name for the mean temperature
                   mean_sal = rep(NA, length(files)), # column name for the mean salinity
                   ) # column name for the year name
cond_data



# write a basic code to calculate the mean and build out
raw_data<-read_csv(paste0(CondPath,"/",files[1])) # test by reading in the first file and see if it works, this is not in a loop yet we are just testing in the i to 1st file works .
head(raw_data)

#test
mean_temp<-mean(raw_data$Temperature, na.rm = TRUE) # calculate a mean
mean_temp

```
### Turn it into a for loop

```{r}

for (i in 1:length(files)){ # loop over 1:3 the number of files

raw_data<-read_csv(paste0(CondPath,"/",files[i]))
#glimpse(raw_data)

cond_data$filename[i]<-files[i]

}

cond_data

```
### Now add in the mean summary
```{r}


for (i in 1:length(files)){ # loop over 1:3 the number of files

raw_data<-read_csv(paste0(CondPath,"/",files[i]))
#glimpse(raw_data)

cond_data$filename[i]<-files[i]
cond_data$mean_temp[i]<-mean(raw_data$Temperature, na.rm =TRUE)
cond_data$mean_sal[i]<-mean(raw_data$Salinity, na.rm =TRUE)
} 

cond_data

```
### {purrr}
purrr cheatsheet

Now, I will teach you to do the exact same thing, but with tidyverse style language.

"purrr enhances R’s functional programming (FP) toolkit by providing a complete and consistent set of tools for working with functions and vectors. If you’ve never heard of FP before, the best place to start is the family of map() functions which allow you to replace many for loops with code that is both more succinct and easier to read. The best place to learn about the map() functions is the iteration chapter in R for data science."

Map functions: The pattern of looping over a vector, doing something to each element and saving the results is so common that the purrr package provides a family of functions to do it for you. There is one function for each type of output:

map() makes a list.
map_lgl() makes a logical vector.
map_int() makes an integer vector.
map_dbl() makes a double vector.
map_chr() makes a character vector.
map_df() makes a dataframe
Each function takes a vector as input, applies a function to each piece, and then returns a new vector that’s the same length (and has the same names) as the input.

### Simple example
There 3 ways to do the same thing in a map() function

Use a canned function that already exists

Let's calculate the mean from a set of random numbers and do it 10 times

Create a vector from 1:10

```{r}
1:10 # a vector from 1 to 10 (we are going to do this 10 times)

# for each time 1:10 make a vector of 15 random numbers based on a normal distribution

1:10 %>% # a vector from 1 to 10 (we are going to do this 10 times) %>% # the vector to iterate over
  map(rnorm, n = 15) # calculate 15 random numbers based on a normal distribution in a list


# Calculate the mean from each list, what is the mean of each 15 values 10 times...

1:10 %>% # a vector from 1 to 10 (we are going to do this 10 times) %>% # the vector to iterate over
  map(rnorm, n = 15)  %>% # calculate 15 random numbers based on a normal distribution in a list 
  map_dbl(mean) # calculate the mean. It is now a vector which is type "double"

# Make your own function, mayby used for gross photosynthesis

1:10 %>% # list 1:10
  map(function(x) rnorm(15, x)) %>% # make your own function, rnorm is the "i"
  map_dbl(mean)


#Use a formula when you want to change the arguments within the function

1:10 %>%
  map(~ rnorm(15, .x)) %>% # changes the arguments inside the function
  map_dbl(mean)
# ~ and .x is like the "i", changing some specific argument within the function


```

### Bring in files using purrr instead of a for loop
Reminder: find the files

```{r}
# point to the location on the computer of the folder
CondPath<-here("Week_13", "data", "cond_data")
files <- dir(path = CondPath,pattern = ".csv")
files

## Or, we can get the full file names in one less step by doing this...

files <- dir(path = CondPath,pattern = ".csv", full.names = TRUE)
#save the entire path name
files

```
# lets read in all the data and stack it on top of each other, then we can mutate to clean it up
```{r}
#read in the files
#Next, read in the files using map instead of a for loop while retaining the filename as a column.

data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") # map everything to a dataframe and put the id in a column called filename
data

## make sure column names are the same across files
```
### calculate means
Now we have a regular dataframe and we can calculate means in the way we already know how! group_by filename and use summarize

```{r}

data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") %>% # map everything to a dataframe and put the id in a column called filename
  group_by(filename) %>%
  summarise(mean_temp = mean(Temperature, na.rm = TRUE),
            mean_sal = mean(Salinity,na.rm = TRUE))
data
```
### Other uses
Maps and loops are awesome for more complicated processes too. For example, you could plot something and save it automatically from each file using similar code.

Read these chapters in R4DS and Advanced R to see all the fancy shortcuts you can do with the purrr package
```{r}
https://r4ds.had.co.nz/iteration.html

https://adv-r.hadley.nz/functionals.html

```

Cool R package of the week
Gganatogram, anatomy diagrams 


