---
title: "week9classwork(week8b)"
author: "Kyle Bosworth"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fi.path = "../outputs/")
ggmap::register_google(key = "AIzaSyDFEiZfRlGQ6VKco6XljqHYg1iuHZBSm3c")
register_stadiamaps ("7d632143-15b1-4847-925d-04842047a2e9", write = TRUE)
```


### Load Libraries
```{r}
library(here)
library(tidyverse)
library(maps)
library(mapdata)
library(mapproj)
library(ggspatial)
library(data.table)
library(ggmap)
library(ggspatial)

library(tidytext)
library(wordcloud2)
library(janeaustenr)

library(patchwork)
library(ggrepel)
library(gganimate)
library(magick)
library(palmerpenguins)

```

Why functions?
- Anytime you have to do a task more then once...
Let's use an example where we want to rescale data in multiple columns (value - min/(max - min))
```{r}
df <- tibble(
  a = rnorm(10), # draws 10 random values from a normal distribution
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
head(df)
```
Anytime you have to do a task more then once...
Let's use an example where we want to rescale data in multiple columns (value - min/(max - min))
Rescale every column individually
```{r}
df<-df %>%
  mutate(a = (a-min(a, na.rm = TRUE))/(max(a, na.rm = TRUE)-min(a, na.rm = TRUE)),
         b = (b-min(b, na.rm = TRUE))/(max(b, na.rm = TRUE)-min(b, na.rm = TRUE)),
         c = (c-min(c, na.rm = TRUE))/(max(c, na.rm = TRUE)-min(c, na.rm = TRUE)),
         d = (d-min(d, na.rm = TRUE))/(max(d, na.rm = TRUE)-min(d, na.rm = TRUE)))

# Now we can write a function for this:
rescale01<- function(x) {
  value<-(x-min(x, na.rm = TRUE))/max(x, na.rm = TRUE)-min(x, na.rm = TRUE)
  return(value)
}

df %>%
  mutate(a = rescale01(a),
         b = rescale01(b),
         c = rescale01(c),
         d = rescale01(d))

```

Breaking it down: there are three key steps to creating a new function:

1. You need to pick a name for the function. Here I’ve used rescale01 because this function rescales a vector to lie between 0 and 1.

2. You list the inputs, or arguments, to the function inside function. Here we have just one argument. If we had more the call would look like function(x, y, z).

3. You place the code you have developed in body of the function, a { block that immediately follows function(...).

4. return tells us what values we want returned when you run the function
```{r}
rescale01 <- function(x) {
  value<-(x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE))
  return(value)
}
```
Note: It is way easier to create a function when you already wrote the code rather than creating the function from scratch.


Letʻs make a function to convert degrees Fahrenheit to Celsius.
 - What is the calculation for F to C?
 
```{r}
temp_C <- (temp_F - 32) * 5 / 9

#1. name the function:

fahrenheit_to_celsius <- function() {
}

#2 put in the equation
fahrenheit_to_celsius <- function() { 
  temp_C <- (temp_F - 32) * 5 / 9
}

#3 decide what the arguments are
fahrenheit_to_celsius <- function(temp_F) {
  temp_C <- (temp_F - 32) * 5 / 9 
}

#4 decide what is being returned
fahrenheit_to_celsius <- function(temp_F) { 
  temp_C <- (temp_F - 32) * 5 / 9 
  return(temp_C)
}

#5 test it
fahrenheit_to_celsius(32)

fahrenheit_to_celsius(212)
```

Making Plots into a function:

Let's say you have a specifically formatted plot that you like and that you plan to use over and over again. By turning it into a function, you only have to code it once.

Let's use our favorite penguin dataset and make a nice plot line plot.

```{r}
library(palmerpenguins)
library(PNWColors) # for the PNW color palette 
# you may need to install the PNWColors library if you haven't used it yet
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
ggplot(penguins, 
       aes(x = body_mass_g, 
           y = bill_length_mm, 
           color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and another example of how to manually change the legend title for colors
  theme_bw()
```
name and set-up the function:
```{r}
#1. 

myplot <- function() {

pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 

ggplot(penguins, aes(x = body_mass_g, y = bill_length_mm, color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()

}

#2 
myplot <- function(data, x, y) {

pal<-pnw_palette("Lake",3, type = "discrete") 

ggplot(penguins, aes(x ={{x}}, y = {{y}}, color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ 
  scale_color_manual("Island", values=pal)+ 
  theme_bw()

}

##TEST it
#before:
  myplot(data = penguins, x = body_mass_g, y = bill_length_mm)
#after: 
  myplot(x = body_mass_g, y = flipper_length_mm)
```
You can also layer onto your plot using '+' just like it is a regular ggplot to change things like labels.
```{r}
myplot(x = body_mass_g, y = flipper_length_mm)+
  labs(x = "Body mass (g)",
       y = "Flipper length (mm)")
```

What are "if-else" statements?
- Imagine you want a variable to be equal to a certain value if a condition is met. This is a typical problem that requires the if ... else ... construct. For instance:
```{r}
a <- 4
b <- 5

# Suppose that if a > b then f should be = to 20, else f should be equal to 10. Using if/else we:

if (a>b) {# my question
  f <- 20 # if it is true, give me answer 1
} else { #else give me answer 2
    f <- 10
}

# test it:
f

```
Plotting using "if-else"
```{r}
# Using if...else... we can make our function even more flexible. Let's say we want the option of adding the geom_smooth lines. We can create a variable that if set to TRUE add the geom_smooth, otherwise print without.

# First add a new argument for lines and make the default TRUE for ease

myplot<-function(data = penguins, x, y ,lines=TRUE ){ # add new argument for lines
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 

ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}

# NOW ADD an "if-else" statement:

myplot<-function(data = penguins, x, y, lines=TRUE ){ # add new argument for lines
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
if(lines==TRUE) {
  ggplot(data, aes(x = {{x}}, 
                   y = {{y}} , 
                   color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
}

myplot(x = body_mass_g, y = flipper_length_mm)

```

