---
title: "week9bclasswork"
author: "Kyle Bosworth"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../outputs/")
knitr::opts_chunk$set(dev = "png", 
                      fig.path = "../outputs/",
                      fig.width = 7,
                      fig.height = 5)
```

### Load Libraries
```{r}
library(here)
library(tidyverse)


```

### Read in data: 
```{r}
#tuesdata <- tidytuesdayR::tt_load(2021, week = 7)
#income_mean<-tuesdata$income_mean

income_mean <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/income_mean.csv')

```

### What is a Factor?

- A specialized version of a character
- used to store categorical data
- The values that a factor takes are called "levels"
    - Default levels are always alphebetical

```{r}
# to make something a factor put the vector in the function facto()
fruits<- factor(c("Apple", "Grape", "Banana"))
fruits
```
### factor booby-traps!
When can this be scary...

Let's say you had a typo in a column of what was suppose to be numbers. R will read everything in as characters. If they are characters and you try to covert it to a number, the rows with real characters will covert to NAs

```{r}
test<-c("A", "1", "2")
as.numeric(test)
```

NA is introduced, factors store everything as integers, factors should only be used used for categorical data, not numerical. 

### {forcats}
The forcats package is a family of functions for categorical data. All the main fuctions start with fct_

Let's start with the starwars dataset to learn about how to use factors
```{r}
glimpse(starwars)
```


### starwars
Let's look at some information by different species in the starwars films. How many individuals of each species are present across all starwars films?

```{r}
# filter out NAs by species

starwars %>%
  filter(!is.na(species)) %>% #count species and sort by TRUE to see how many unique species there are and how many characters there are
  count(species)

```
There are 38 unique species, but most are really rare. Let's say we wanted to lump all the species together that had less than 3 individuals. We could use fct_lump() which converts the data into a factor and lumps it together

```{r}
star_counts<- starwars %>%
  filter(!is.na(species)) %>%
  mutate(species = fct_lump(species, n = 3)) %>% #groups species that occur less than 3 into a new category
  count(species)
```

### reordering factors
A basic ggplot
```{r}
star_counts %>%
  ggplot(aes(x = fct_reorder(species, n, .desc = TRUE), y = n))+ # reorder the factor of species by n
  geom_col() +
  labs(x = "Species")
```
### Reordering line plots
Let's say you wanted to make a line plot and reorder the legend to match the order of the lines. We will use the income_mean dataset to illustrate this.

```{r}
glimpse(income_mean)
```
How is income changing over time? We will make a plot of the total income by year and quantile across all dollar types.

```{r}
total_income<-income_mean %>%
  group_by(year, income_quintile) %>%
  summarise(income_dollars_sum = sum(income_dollars)) %>%
  mutate(income_quintile = factor(income_quintile)) # make it a factor
```
Reordering Plots! With data that already exists
Basic plot:
```{r}
total_income %>%
ggplot(aes(x = year,
             y = income_dollars_sum,
             color = fct_reorder2(income_quintile, year, income_dollars_sum)))+ #reordered by incomequintile column, year, then sum
  geom_line() +
labs(color = "income quantile") 
```

Reordering levels directly in a vector just cause

```{r}
x1 <- factor(c("Jan", "Mar", "Apr","Dec "))
x1

x1 <- factor(c("Jan", "Mar", "Apr", "Dec"), levels = c("Jan", "Mar", "Apr", "Dec")) #within factor function you can specify the specific order that we want! factor, then levels of that factor
x1
```
### Subset data with factors
Back to starwars. Let's say we instead of grouping our species that have < 3 counts into "Other" we just filter them out.

```{r}
starwars_clean<-starwars %>% 
  filter(!is.na(species)) %>% # remove the NAs
  count(species, sort = TRUE) %>%
  mutate(species = factor(species)) %>% #makes species into factor
  filter(n>3) %>% #only keep species that have more than 3
  droplevels() %>% # drop extra levels
  mutate(species = fct_recode(species, "Humanoid" = "Human")) #recode/rename your factors by using fct recode, "new name" = "old name"
  
  
  starwars_clean
levels(starwars_clean)

```
Totally awsome R package: 
Today's totally awesome R package
{gm} (generate music)


```{r}
install.packages('gm')
```

Then follow directions from here to download the MuseScore program and tell R where the path to that program is.


```{r}
library(gm)
m <- 
  Music() +  # initialize a Music object
  Meter(4, 4) + # add a 4/4 time signature
  Line(c("C5", "D5", "E5", "F5")) # add a musical line of four quarter notes
gm::show(m, to = c("score", "audio"))
```


