---
title: "Tidy Tuesday 4"
author: "Kyle Bosworth"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../outputs/")
knitr::opts_chunk$set(dev = "png", 
                      fig.path = "../outputs/",
                      fig.width = 7,
                      fig.height = 5)
```

### Load Libraries
```{r}
library(here)
library(tidyverse)
library(plotly)
library(dplyr)
library(ggbiplot) # specific package for PCA visualization

```

### Read in Data from TidyTuesday GitHub
```{r}
episode_metrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-11-19/episode_metrics.csv')

```

Goal: Create a PCA Plot
I have wanted to learn how to create a PCA plot for this tidy tuesday I will be using the Bobs Burgers Data set. I thought that this would be a cool dataset to try this out on because of the type of values (numerical) are in each category of this data set. 

### Prepare the data
```{r}


# Filter data so that we only include the seasons 1-3 first, using all seasons because too much too look at all at once
episode_metrics_filtered <- episode_metrics %>%
  filter(season %in% c(1, 2, 3))

# select the variables for PCA
pcadata <- episode_metrics_filtered %>%
  select(dialogue_density, unique_words, exclamation_ratio) %>%
  na.omit() %>%
  scale() # scale() standardizes the variable to have a mean = 0 and sd = 1, important for PCA

  
```

### Preform PCA
```{r}
# using the scaled data, preform PCA using prcomp
PCAresults <- prcomp(pcadata, scale. = TRUE)

# calculate the proportion of variance explained by each individal Principal Component (PC)
var_explained <- PCAresults$sdev^2 / sum(PCAresults$sdev^2) * 100

```

### Creating the PCA biplot using ggbiplot
```{r}

PCAplot <- ggbiplot(PCAresults, 
                    obs.scale = 1, #scales
                    var.scale = 1,
                    groups = factor(episode_metrics_filtered$season), #colorby filtered seasons
                    ellipse = TRUE, # this adds confidence eclipses
                    circle = FALSE) + #removed confidence circles to simplify the plot
  xlim(-4, 4) +  # these dictate the x-axis limits
  ylim(-4, 6) +  # y-axis limits
  scale_color_manual(name = "Season",
                    values = c("orange", "lightblue", "forestgreen"),
                    labels = c("Season 1", "Season 2", "Season 3")) +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.justification = c(1, 1),
    legend.box.background = element_rect(color = "gray", fill = "white", size = 0.5),
    legend.margin = margin(6, 6, 6, 6),
    panel.grid.major = element_line(color = "gray"),
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10)
  ) +
  labs(
    title = "PCA of Bob's Burgers Episodes (Seasons 1-3)",
    x = sprintf("PC1 (%.1f%% explained var.)", var_explained[1]), #adds titles to inform variance trends
    y = sprintf("PC2 (%.1f%% explained var.)", var_explained[2])
  )
print(PCAplot)
```



PC1 - 42.1% of the variance - represent the direction of maximum variation in this simplified data set (Seasons 1-3). It shows the strongest pattern in how dialog density, unique words, and exclamation ratio vary together across episodes/seasons.

PC2 - 35.6% of the variance - represents the second most significant pattern of variation, uncorrelated to PC1. 

Both PC1 and PC2 explain 77.7% of the dataset. We can notice that The show has been evolving in terms of script depth. We can see that the dialog density and unique words point in the same direction (positive correlation) and increase from season 1 to season 3. 